---
title: "`tinyjpg()` Benchmarks"
output:
  html:
    meta:
      css: ["@tabsets"]
      js: ["@tabsets"]
---

```{r}
litedown::reactor(echo = FALSE)
library(tinyimg)
options(xfun.md_table.limit = Inf)
```

## Overview

This document benchmarks the `tinyjpg()` function from the **tinyimg** package
across different quality levels (50–100). We test with three types of plots:

1. A scatterplot
2. A simple boxplot
3. A complex 3D perspective plot

For each quality level, we measure file size reduction (in KB and percentage)
and time taken for optimization.

## Test Plots

We create test JPEG files at high quality (95) to serve as originals:

```{r scatterplot, fig.cap='Flipper length vs bill length of 344 penguins.'}
par(mar = c(4, 4, 1, 1))
n = length(penguins$flipper_len)
cols = grDevices::hcl.colors(n, "Spectral")
plot(penguins$flipper_len, penguins$bill_len, pch = 19,
     col = cols[seq_along(penguins$flipper_len)],
     xlab = "Flipper Length (mm)", ylab = "Bill Length (mm)")
```

```{r boxplot, fig.cap='Miles per gallon by cylinder count.'}
par(mar = c(4, 4, 1, 1))
boxplot(mpg ~ cyl, data = mtcars, xlab = "Number of Cylinders",
        ylab = "Miles Per Gallon",
        col = c("lightblue", "lightgreen", "lightpink"))
```

```{r persp, fig.cap='A complex 3D surface defined by the function sin(r)/r.'}
par(mar = c(2, 2, 1, 1))
x = seq(-10, 10, length = 50); y = seq(-10, 10, length = 50)
z = outer(x, y, function(x, y) { r = sqrt(x^2 + y^2); 10 * sin(r) / r })
z[is.nan(z)] = 10
persp(x, y, z, theta = 30, phi = 20, col = "lightblue", shade = 0.4,
      xlab = "X", ylab = "Y", zlab = "Z")
```

```{r create-jpgs}
# Save the above plots as JPEG files (quality 95) to use as originals
png_files = litedown::get_context("plot_files")

create_jpeg = function(input_png, quality = 95) {
  out = sub("\\.png$", ".jpg", input_png)
  img = png::readPNG(input_png)
  jpeg(out, width = dim(img)[2], height = dim(img)[1], quality = quality,
       units = "px")
  graphics::rasterImage(img, 0, 0, 1, 1, xlim = c(0, 1), ylim = c(0, 1))
  dev.off()
  out
}

# Fall back: write JPEG directly if png package unavailable
make_orig_jpg = function(code_expr, nm) {
  f = tempfile(fileext = ".jpg")
  jpeg(f, width = 800, height = 600, quality = 95)
  eval(code_expr)
  dev.off()
  f
}

scatter_orig = make_orig_jpg(quote({
  par(mar = c(4, 4, 1, 1))
  n = length(penguins$flipper_len)
  cols = grDevices::hcl.colors(n, "Spectral")
  plot(penguins$flipper_len, penguins$bill_len, pch = 19,
       col = cols[seq_along(penguins$flipper_len)],
       xlab = "Flipper Length (mm)", ylab = "Bill Length (mm)")
}), "scatter")

boxplot_orig = make_orig_jpg(quote({
  par(mar = c(4, 4, 1, 1))
  boxplot(mpg ~ cyl, data = mtcars, xlab = "Number of Cylinders",
          ylab = "Miles Per Gallon",
          col = c("lightblue", "lightgreen", "lightpink"))
}), "boxplot")

persp_orig = make_orig_jpg(quote({
  par(mar = c(2, 2, 1, 1))
  x = seq(-10, 10, length = 50); y = seq(-10, 10, length = 50)
  z = outer(x, y, function(x, y) { r = sqrt(x^2 + y^2); 10 * sin(r) / r })
  z[is.nan(z)] = 10
  persp(x, y, z, theta = 30, phi = 20, col = "lightblue", shade = 0.4,
        xlab = "X", ylab = "Y", zlab = "Z")
}), "persp")

orig_files = c(scatter = scatter_orig, boxplot = boxplot_orig, persp = persp_orig)
```

## Benchmarks

Now let's optimize these JPEG files at different quality levels and measure
the results:

```{r benchmark}
quality_levels = c(50, 60, 70, 75, 80, 90, 95, 100)

benchmark_file = function(input_file, plot_name) {
  original_size = file.info(input_file)$size

  raw_row = data.frame(
    Plot = plot_name, Quality = NA_integer_,
    Size_KB = original_size / 1024, Reduction_pct = 0, Time_sec = 0
  )

  opt_rows = do.call(rbind, lapply(quality_levels, function(q) {
    out = tempfile(fileext = ".jpg")
    t = system.time(tinyjpg(input_file, out, quality = q, verbose = FALSE))["elapsed"]
    new_size = file.info(out)$size
    data.frame(
      Plot = plot_name, Quality = q,
      Size_KB = new_size / 1024,
      Reduction_pct = 100 * (original_size - new_size) / original_size,
      Time_sec = t
    )
  }))

  rbind(raw_row, opt_rows)
}

results = rbind(
  benchmark_file(orig_files["scatter"], "scatterplot"),
  benchmark_file(orig_files["boxplot"],  "boxplot"),
  benchmark_file(orig_files["persp"],    "persp")
)

results$Size_KB       = round(results$Size_KB, 1)
results$Reduction_pct = round(results$Reduction_pct, 1)
results$Time_sec      = round(results$Time_sec, 3)
```

### Results table

::: tabset

#### scatterplot

```{r scatter-table}
knitr::kable(
  results[results$Plot == "scatterplot", -1],
  row.names = FALSE,
  col.names = c("Quality", "Size (KB)", "Reduction (%)", "Time (s)")
)
```

#### boxplot

```{r boxplot-table}
knitr::kable(
  results[results$Plot == "boxplot", -1],
  row.names = FALSE,
  col.names = c("Quality", "Size (KB)", "Reduction (%)", "Time (s)")
)
```

#### persp

```{r persp-table}
knitr::kable(
  results[results$Plot == "persp", -1],
  row.names = FALSE,
  col.names = c("Quality", "Size (KB)", "Reduction (%)", "Time (s)")
)
```

:::

## Visualization

### File size by quality

```{r size-plot, fig.cap='File size as a percentage of the original across quality levels.'}
par(mar = c(4, 4, 1, 1))
cols = c(scatterplot = "steelblue", boxplot = "tomato", persp = "seagreen")
plot(range(quality_levels), c(0, 120), type = "n",
     xlab = "Quality", ylab = "File size (% of original)")
abline(h = 100, lty = 2, col = "grey70")
for (nm in names(cols)) {
  sub = results[results$Plot == nm & !is.na(results$Quality), ]
  pct = 100 - sub$Reduction_pct
  lines(sub$Quality, pct, col = cols[nm], lwd = 2)
  points(sub$Quality, pct, col = cols[nm], pch = 19)
}
legend("topleft", names(cols), col = cols, lwd = 2, pch = 19, bty = "n")
```

### Optimization time

```{r time-plot, fig.cap='Optimization time across quality levels.'}
par(mar = c(4, 4, 1, 1))
plot(range(quality_levels), c(0, max(results$Time_sec, na.rm = TRUE) * 1.1),
     type = "n", xlab = "Quality", ylab = "Time (seconds)")
for (nm in names(cols)) {
  sub = results[results$Plot == nm & !is.na(results$Quality), ]
  lines(sub$Quality, sub$Time_sec, col = cols[nm], lwd = 2)
  points(sub$Quality, sub$Time_sec, col = cols[nm], pch = 19)
}
legend("topleft", names(cols), col = cols, lwd = 2, pch = 19, bty = "n")
```

## Summary

Average results across the three plots:

```{r summary}
agg = aggregate(cbind(Reduction_pct, Time_sec) ~ Quality,
                data = results[!is.na(results$Quality), ], FUN = mean)
agg$Reduction_pct = round(agg$Reduction_pct, 1)
agg$Time_sec      = round(agg$Time_sec, 3)
names(agg) = c("Quality", "Mean reduction (%)", "Mean time (s)")
knitr::kable(agg, row.names = FALSE)
```
