---
title: "`tinyjpg()` Benchmarks"
output:
  html:
    meta:
      css: ["@tabsets"]
      js: ["@tabsets"]
---

```{r}
litedown::reactor(echo = FALSE, dev = 'jpeg')
library(tinyimg)
options(xfun.md_table.limit = Inf)
```

## Overview

This document benchmarks the `tinyjpg()` function from the **tinyimg** package
across different quality levels (50–100). We test with three types of plots:

1. A scatterplot based on the penguins dataset
2. A simple boxplot (simpler graphics)
3. A complex 3D perspective plot (more complex graphics)

For each quality level, we measure:

- File size reduction (in KB and percentage)
- Time taken for optimization

## Test Plots

First, we create test plots using the `jpeg()` device (set as default via
`dev = 'jpeg'`):

```{r scatterplot, fig.cap='Penguin bill length vs flipper length.'}
par(mar = c(4, 4, 1, 1))
n = length(penguins$flipper_len)
cols = grDevices::hcl.colors(n, "Spectral")
plot(penguins$flipper_len, penguins$bill_len,
     pch = 19,
     col = cols[seq_along(penguins$flipper_len)],
     xlab = "Flipper Length (mm)",
     ylab = "Bill Length (mm)")
```

```{r boxplot, fig.cap='Miles per gallon by cylinder count.'}
par(mar = c(4, 4, 1, 1))
boxplot(mpg ~ cyl, data = mtcars,
        xlab = "Number of Cylinders",
        ylab = "Miles Per Gallon",
        col = c("lightblue", "lightgreen", "lightpink"))
```

```{r persp, fig.cap='A complex 3D surface defined by the function sin(r)/r.'}
par(mar = c(2, 2, 1, 1))
x = seq(-10, 10, length = 50)
y = seq(-10, 10, length = 50)
z = outer(x, y, function(x, y) {
  r = sqrt(x^2 + y^2)
  10 * sin(r) / r
})
z[is.nan(z)] = 10
persp(x, y, z,
      theta = 30, phi = 30,
      expand = 0.5,
      col = "lightblue",
      shade = 0.5,
      ticktype = "detailed",
      xlab = "X", ylab = "Y", zlab = "Z")
```

## Benchmarks

Now let's optimize these JPEG files at different quality levels and measure
the results:

```{r benchmark}
# Get the JPEG plot files created by litedown
plot_files = litedown::get_context("plot_files")

# Create a temporary directory for optimized files
tmp_dir = tempfile()
dir.create(tmp_dir)

# Function to benchmark a single file at all quality levels
benchmark_file = function(input_file, plot_name) {
  original_size = file.info(input_file)$size

  # Add a row for the original file (quality = NA)
  raw_row = data.frame(
    Plot = plot_name,
    Quality = NA_integer_,
    Size_KB = original_size / 1024,
    Reduction_pct = 0,
    Time_sec = 0
  )

  optimized_rows = do.call(rbind, lapply(c(50, 60, 70, 75, 80, 90, 95, 100), function(q) {
    output_file = file.path(tmp_dir, sprintf("%s_q%d.jpg", plot_name, q))

    time_taken = system.time({
      tinyjpg(input_file, output = output_file, quality = q)
    })["elapsed"]

    new_size = file.info(output_file)$size
    reduction_pct = ((original_size - new_size) / original_size) * 100

    data.frame(
      Plot = plot_name,
      Quality = q,
      Size_KB = new_size / 1024,
      Reduction_pct = reduction_pct,
      Time_sec = time_taken
    )
  }))

  rbind(raw_row, optimized_rows)
}

# Benchmark all three plots
results = rbind(
  benchmark_file(plot_files[1], "scatterplot"),
  benchmark_file(plot_files[2], "boxplot"),
  benchmark_file(plot_files[3], "persp")
)

rownames(results) = NULL
results
```

## Visualization

```{r plot-filesize, fig.cap='File size vs quality level.'}
par(mar = c(4, 4, 1, 1))
plot_names = unique(results$Plot)
pchs = c(19, 17, 15)
quality_levels = c(50, 60, 70, 75, 80, 90, 95, 100)
opt_results = subset(results, !is.na(Quality))

plot(NA, xlim = range(quality_levels), ylim = range(opt_results$Size_KB),
     xlab = "Quality", ylab = "File Size (KB)")
for (i in seq_along(plot_names)) {
  plot_data = subset(opt_results, Plot == plot_names[i])
  lines(plot_data$Quality, plot_data$Size_KB,
        type = "b", pch = pchs[i], col = i + 1)
}
legend("topleft", legend = plot_names,
       col = seq_along(plot_names) + 1,
       pch = pchs[seq_along(plot_names)], lty = 1)
grid()
```

```{r plot-reduction, fig.cap='Size reduction percentage vs quality level.'}
par(mar = c(4, 4, 1, 1))
plot(NA, xlim = range(quality_levels), ylim = range(opt_results$Reduction_pct),
     xlab = "Quality", ylab = "Size Reduction (%)")
for (i in seq_along(plot_names)) {
  plot_data = subset(opt_results, Plot == plot_names[i])
  lines(plot_data$Quality, plot_data$Reduction_pct,
        type = "b", pch = pchs[i], col = i + 1)
}
legend("topright", legend = plot_names,
       col = seq_along(plot_names) + 1,
       pch = pchs[seq_along(plot_names)], lty = 1)
grid()
```

```{r plot-time, fig.cap='Processing time vs quality level.'}
par(mar = c(4, 4, 1, 1))
plot(NA, xlim = range(quality_levels), ylim = range(opt_results$Time_sec),
     xlab = "Quality", ylab = "Time (seconds)")
for (i in seq_along(plot_names)) {
  plot_data = subset(opt_results, Plot == plot_names[i])
  lines(plot_data$Quality, plot_data$Time_sec,
        type = "b", pch = pchs[i], col = i + 1)
}
legend("topleft", legend = plot_names,
       col = seq_along(plot_names) + 1,
       pch = pchs[seq_along(plot_names)], lty = 1)
grid()
```

```{r optimize-document-plots, include=FALSE}
tinyjpg(litedown::get_context("plot_files"), quality = 75)
```
