---
title: "`tinypng()` Benchmarks"
---

```{r}
litedown::reactor(echo = FALSE)
library(tinyimg)
options(xfun.md_table.limit = Inf)
```

## Overview

This document benchmarks the `tinypng()` function from the **tinyimg** package
across different optimization levels (0-6). We test with three types of plots:

1. A scatterplot based on the penguins dataset
2. A simple boxplot (simpler graphics)
3. A complex 3D perspective plot (more complex graphics)

For each optimization level, we measure:

- File size reduction (in KB and percentage)
- Time taken for optimization

## Test Plots

First, we create test plots using the `png()` device:

```{r create-penguins, fig.cap='Penguin bill length vs flipper length.'}
par(mar = c(4, 4, 1, 1))
plot(penguins$flipper_len, penguins$bill_len,
     pch = 19,
     xlab = "Flipper Length (mm)",
     ylab = "Bill Length (mm)")
```

```{r create-boxplot, fig.cap='Miles per gallon by cylinder count.'}
par(mar = c(4, 4, 1, 1))
boxplot(mpg ~ cyl, data = mtcars, 
        xlab = "Number of Cylinders", 
        ylab = "Miles Per Gallon",
        col = c("lightblue", "lightgreen", "lightpink"))
```

```{r create-persp, fig.cap='A complex 3D surface defined by the function sin(r)/r.'}
par(mar = c(2, 2, 1, 1))
x = seq(-10, 10, length = 50)
y = seq(-10, 10, length = 50)
z = outer(x, y, function(x, y) {
  r = sqrt(x^2 + y^2)
  10 * sin(r) / r
})

persp(x, y, z, 
      theta = 30, phi = 30, 
      expand = 0.5,
      col = "lightblue",
      shade = 0.5,
      ticktype = "detailed",
      xlab = "X", ylab = "Y", zlab = "Z")
```

## Benchmarks

Now let's optimize these plot files at different levels and measure the results:

```{r benchmark}
# Get the plot files created by litedown
plot_files = litedown::get_context("plot_files")

# Create a temporary directory for optimized files
tmp_dir = tempfile()
dir.create(tmp_dir)

# Function to benchmark a single file at all levels
benchmark_file = function(input_file, plot_name) {
  original_size = file.info(input_file)$size
  
  # Add a row for the raw file (level = -1)
  raw_row = data.frame(
    Plot = plot_name,
    Level = -1L,
    Size_KB = original_size / 1024,
    Reduction_pct = 0,
    Time_sec = 0
  )
  
  optimized_rows = do.call(rbind, lapply(0:6, function(level) {
    output_file = file.path(tmp_dir, sprintf("%s_level_%d.png", plot_name, level))
    
    time_taken = system.time({
      tinypng(input_file, output = output_file, level = level)
    })["elapsed"]
    
    new_size = file.info(output_file)$size
    reduction_pct = ((original_size - new_size) / original_size) * 100
    
    data.frame(
      Plot = plot_name,
      Level = level,
      Size_KB = new_size / 1024,
      Reduction_pct = reduction_pct,
      Time_sec = time_taken
    )
  }))
  
  rbind(raw_row, optimized_rows)
}

# Benchmark all three plots
results = rbind(
  benchmark_file(plot_files[1], "Scatterplot"),
  benchmark_file(plot_files[2], "Boxplot"),
  benchmark_file(plot_files[3], "Persp")
)

rownames(results) = NULL
results
```

Note: The optimization level of `-1` represents the raw, unoptimized file (not a valid value for the `level` argument of `tinypng()`). This helps show the effect of `level = 0` more clearly.

## Lossy Optimization

```{r lossy-benchmark}
lossy_levels = c(0, 0.25, 0.5, 0.75, 1)
plot_names = c("Scatterplot", "Boxplot", "Persp")
lossy_files = list()
lossy_rows = list()

for (i in seq_along(plot_files)) {
  input_file = plot_files[i]
  plot_name = plot_names[i]
  original_size = file.info(input_file)$size
  lossy_files[[plot_name]] = list()
  for (lossy in lossy_levels) {
    output_file = file.path(
      dirname(input_file), sprintf("%s_lossy_%s.png", plot_name, gsub("\\.", "_", lossy))
    )
    time_taken = system.time({
      tinypng(input_file, output = output_file, lossy = lossy)
    })["elapsed"]
    lossy_files[[plot_name]][[as.character(lossy)]] = output_file
    new_size = file.info(output_file)$size
    lossy_rows[[length(lossy_rows) + 1L]] = data.frame(
      Plot = plot_name,
      Lossy = lossy,
      Size_KB = new_size / 1024,
      Reduction_pct = ((original_size - new_size) / original_size) * 100,
      Time_sec = time_taken
    )
  }
}
lossy_results = do.call(rbind, lossy_rows)

rownames(lossy_results) = NULL
lossy_results
```

### Visual comparison {.tabset}

```{r lossy-images, results='asis'}
for (plot_name in names(lossy_files)) {
  cat("#### ", plot_name, " {.tabset}\n\n", sep = "")
  for (lossy in names(lossy_files[[plot_name]])) {
    img = lossy_files[[plot_name]][[lossy]]
    cat("##### lossy = ", lossy, "\n\n", sep = "")
    cat("![](", img, ")\n\n", sep = "")
  }
}
```

## Visualization

```{r plot-filesize, fig.cap='File size vs optimization level (log scale).'}
par(mar = c(4, 4, 1, 1))

# Prepare data for all plots
plot_names = unique(results$Plot)
pchs = c(19, 17, 15)

# Calculate y-axis range on log scale
all_sizes = results$Size_KB
ylim_range = range(all_sizes)

plot(NA, xlim = c(-1, 6), ylim = ylim_range, log = "y",
     xlab = "Optimization Level", 
     ylab = "File Size (KB, log scale)")
for (i in seq_along(plot_names)) {
  plot_data = subset(results, Plot == plot_names[i])
  lines(plot_data$Level, plot_data$Size_KB, 
        type = "b", pch = pchs[i], col = i + 1)
}
legend("topright", legend = plot_names, 
       col = seq_along(plot_names) + 1, 
       pch = pchs[seq_along(plot_names)], lty = 1)
grid()
```

```{r plot-time, fig.cap='Processing time vs optimization level.'}
par(mar = c(4, 4, 1, 1))

# Calculate y-axis range
all_times = results$Time_sec
ylim_range = range(all_times)

plot(NA, xlim = c(-1, 6), ylim = ylim_range,
     xlab = "Optimization Level", 
     ylab = "Time (seconds)")
for (i in seq_along(plot_names)) {
  plot_data = subset(results, Plot == plot_names[i])
  lines(plot_data$Level, plot_data$Time_sec, 
        type = "b", pch = pchs[i], col = i + 1)
}
legend("topleft", legend = plot_names, 
       col = seq_along(plot_names) + 1, 
       pch = pchs[seq_along(plot_names)], lty = 1)
grid()
```

```{r plot-reduction-pct, fig.cap='Size reduction percentage by optimization level.'}
par(mar = c(4, 4, 1, 1))

# Calculate y-axis range
all_reduction = results$Reduction_pct
ylim_range = range(all_reduction)

plot(NA, xlim = c(-1, 6), ylim = ylim_range,
     xlab = "Optimization Level", 
     ylab = "Size Reduction (%)")
for (i in seq_along(plot_names)) {
  plot_data = subset(results, Plot == plot_names[i])
  lines(plot_data$Level, plot_data$Reduction_pct, 
        type = "b", pch = pchs[i], col = i + 1)
}
legend("bottomright", legend = plot_names, 
       col = seq_along(plot_names) + 1, 
       pch = pchs[seq_along(plot_names)], lty = 1)
grid()
```

```{r plot-size-reduction-bar, fig.cap='Absolute size reduction by optimization level.'}
par(mar = c(4, 4, 1, 1))

# Prepare data for bar chart - reduction from raw (level -1) to each level
reduction_matrix = do.call(cbind, lapply(plot_names, function(pn) {
  plot_data = subset(results, Plot == pn)
  raw_size = plot_data$Size_KB[plot_data$Level == -1]
  optimized_sizes = plot_data$Size_KB[plot_data$Level >= 0]
  raw_size - optimized_sizes
}))

barplot(t(reduction_matrix),
        beside = TRUE,
        names.arg = 0:6,
        col = seq_along(plot_names) + 1,
        xlab = "Optimization Level",
        ylab = "Size Reduction (KB)",
        legend.text = plot_names,
        args.legend = list(x = "topleft"))
grid()
```

```{r optimize-document-plots, include=FALSE}
tinypng(litedown::get_context("plot_files"), level = 6)
```

```{r cleanup, include=FALSE}
unlink(tmp_dir, recursive = TRUE)
```
